Type-Safe Modular Parsing

### 2017-01-11 paper/sections/Inheritance.tex
Some revision for S4
Fix 3.4

Algebraic datatypes and grammars fit well together. Indeed there are
obvious similarities in declaring a grammar, and declaring an
algebraic datatype. It is difficult to add new language
constructs to both grammars and ASTs based on algebraic datatypes. However,
ASTs modelled with algebraic datatypes naturally allow for the easy
addition of operations. So using such ASTs at least preserves one
dimension of extensibility.

@@

\subsection{Multiple Extensions and Independent Extensibility}
A nice feature of Scala is its support for a form of multiple inheritance.
Scala has a linearized-style multiple inheritance for traits~\cite{}. 
This can be very helpful when composing several languages, which 
may have been independently developed. Suppose now
we want to compose the parsers for expressions from pre-defined
languages \inlinecode{LanguageA} and \inlinecode{LanguageB} using
alternative.\bruno{Don't be abstract! Use concrete examples. You 
already have Var, maybe you can add another language extension for 
boolean literals. Show the code for the boolean literals parsing as
well as the composition code.}
The use of keyword \inlinecode{super} is able to
specify the implementation from inheritance, like:

\lstinputlisting[linerange=70-76]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE

Interestingly note that, due to the use of multiple inheritance, we 
need two different super calls.\bruno{expand here. People may not 
be familiar with scala super calls, you have to explain what the 
syntax does.}
As we demonstrated, inheritance is the key technique to obtain semantic modularity.
It enables type-safe code reuse and separate compilation for parsing OO style ASTs.

>>>>

Algebraic datatypes and grammars fit well together. Indeed there are
obvious similarities in declaring a grammar, and declaring an
algebraic datatype. It is difficult to modularly add new language
constructs to both grammars and ASTs based on algebraic datatypes. However,
ASTs modelled with algebraic datatypes naturally allow for the easy
addition of operations. So using such ASTs at least preserves one
dimension of extensibility.

@@

\subsection{Multiple Extensions and Independent Extensibility}
A nice feature of Scala is its support for a form of multiple inheritance.
Scala has a linearized-style multiple inheritance for traits~\cite{}.
This can be very helpful when composing several languages, which
may have been independently developed. Suppose now
we want to compose the parsers for expressions from pre-defined
languages \inlinecode{LanguageA} and \inlinecode{LanguageB} using
alternative. The new parser can be built by inheriting both parsers
at the same time:

\lstinputlisting[linerange=67-73]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE

\huang{added explanation of the `super' keyword}

The \inlinecode{super[T].x} syntax in Scala, so called \textit{static super reference}, refers to the type or method \inlinecode{x} in the parent trait \inlinecode{T}. Under multiple inheritance, it can be used to distinguish the method of the same name. Therefore in the new parser, we use \inlinecode{super} to specify which \inlinecode{pExpr} we are referring to.

\huang{added a concrete example}

As a concrete example, we compose the language of integers and additions with a language of boolean values and if-then-else expressions, to build a new language.

\lstinputlisting[linerange=79-95]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=BOOL_EXAMPLE

\begin{comment}
\bruno{You can present the abstract example, as you do
  here, but you should also present a concrete example. You
already have Var, maybe you can add another language extension for
boolean literals. Show the code for the boolean literals parsing as
well as the composition code.}

Interestingly note that, due to the use of multiple inheritance, we
need two different super calls.\bruno{expand here. People may not
be familiar with scala super calls, you have to explain what the
syntax does.}
\end{comment}

Then the new parser, which contains both integer and boolean expressions, is able to parse hybrid expressions such as \inlinecode{"if true then 1 + 2 else 3 + 4"}.

As we demonstrated, inheritance with method overriding is the key technique to obtain semantic modularity.
It enables type-safe code reuse and separate compilation for parsing OO style ASTs.
